---
title: "Variational Bayes Principal Components Analysis"
author: "Tuan Nguyen"
date: "4/23/2020"
output: 
  pdf_document: 
    keep_tex: yes
    latex_engine: xelatex
---
# Create a problem, implement the Principal Components Analysis (PCA) Variational Bayes inference technique.
# Import package Rdimtools, an R suite of a number of dimension reduction and estimation methods.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rdimtools)
```
# Generate helix-shaped data
```{r}
X <- aux.gensamples(n=160, noise=0.01, dname='helix')
```
# Run PCA and VBPCA inferences
```{r}
pca  <- do.pca(X, ndim=2, preprocess='center')
vbpca  <- do.bpca(X, ndim=2, preprocess='center')
```
# Compare PCA and VBPCA by visualization
```{r}
opar <- par(no.readonly=TRUE)
par(mfrow=c(1,2))
plot(pca$Y, main='PCA')
plot(vbpca$Y, main='VBPCA')
par(opar)
```
# Print out parameter of the VBPCA algorithm
```{r}
cat('The algorithm converges after', vbpca$mp.itercount, 'iterations.\n')
cat('Estimated squared sigma value:', vbpca$mp.sigma2, '.\n')
```
# The helix-shaped distribution of the posterior means of the latent variables is in identical shape of the standard PCA projections after 7 iterations.